---
title: "Chapter 4"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r, message=FALSE}
library(magrittr)
library(tidyverse)
library(dplyr)
library(psych)
library(smacof)
library(ggplot2)
library(cmdpref)
```

# Example 1 (Airlines Data)

* Data is distances between pairs of US cities
* Use MDS to obtain map of cities
* Small values are similar

```{r}
airlines<- read.csv("~/MSTAT/STAT8019_Marketing_Analytics/lecture notes/chap4r/airlines.csv", row.names = 1, header = T)

airlines
```

**Solution**

* Attribute free (only distance)

* Transformation
  * Ratio is meaningful
  * Other transformation is possible

## Fit MDS

* Can vary the number of dimensions
* Small similar, large difference
* IF the other way round, need to convert, maybe use reciprocal

```{r}
#mds() multidimensional scaling
#delta= dissimilarity matrix
#ndim= number of dimensions
#type="ratio" ratio transformation
fit1<-mds(delta=airlines,ndim=1,type="ratio")
fit1
fit2<-mds(delta=airlines,ndim=2,type="ratio")
fit2
fit3<-mds(delta=airlines,ndim=3,type="ratio")
fit3
fit4<-mds(delta=airlines,ndim=4,type="ratio")
fit4
```

### Stress Scree Plot

```{r}
#scree plot
stress<-data.frame(dim=1:4,stress=c(fit1$stress,fit2$stress,fit3$stress,fit4$stress))
ggplot(data=stress,aes(x=dim,y=stress))+
  geom_line()+
  geom_point()
```

* 2 dimensins are the best, lowest stress
* Not necessarily continue decreasing
* Similar to map, 2 dimensions are enough

## Coordinates

```{r}
#configuration
summary(fit2)
```

* Flip xy axis to flip for better illustration

```{r}
#flip horizontally and vertically so as to be consistent with US map
fit2$conf<- -fit2$conf
#plot the configuration
plot(fit2,xlim=c(-1.5,1.2),ylim=c(-0.5,0.7),main="",cex=2)
```

### Divided (Ratio) Distance from coordinates

* Will return the ratio distance rather than the original absolute distance

```{r}
#derived distances
fit2$confdist
```

```{r}
#derived distances vs original distances
plot(fit2,plot.type="Shepard")
```

* Straight line implies what we observe and what we divide are consistent with each other

# Example 2 (Cars Data)

* ranking of 55 pairs
* 1 - 55
* Smaller rank is similar, larger rank is dissimilar


```{r}
cars<-read.csv("~/MSTAT/STAT8019_Marketing_Analytics/lecture notes/chap4r/cars.csv", header = TRUE,row.names=1)
cars
```

## `"interval"` Transformation

* D* = a + bD

```{r}
fit <- mds(cars, ndim = 2, type="interval")
summary(fit)
```

```{r}
fit$stress
```

### Perceptual Plot

```{r}
plot(fit,xlim=c(-1.2,1.3),main="",cex=2)
```


## `"ordinal"` transformation

* Ordering of the data is meaningful
* Others such as difference, ratio, are not meaningful

```{r}
fit2 <- mds(cars, ndim = 2, type="ordinal")
summary(fit2)
```

```{r}
fit2$stress
```

* Clearly it is the ordinal level analysis that fits the data better

```{r}
plot(fit2,xlim=c(-1.2,1.3),main="",cex=1)
```


# Example 3 (Newspaper data)

* Using `mdpref` data because data structure is preference type
* Larger number is higher preference, lower is lesser preference

```{r}
newspaper<-read.csv("~/MSTAT/STAT8019_Marketing_Analytics/lecture notes/chap4r/newspaper1t.csv", header = TRUE,row.names=1)
newspaper
```

## `cmdpref` model

```{r}
fit<-cmdpref(newspaper,monotone=T)

fit
```

```{r}
summary(fit)
```

## Perceptual Map

```{r}
par(mar=c(2,1,2,1))
plot(fit)
```

* Dimension 1 splits by global vs local
  * global - NYT, BG
  * local - BH, NYP, NYDW
  
* Dimension 2 , some difference between BH, NYP, BG vs NYT, NYDN

* Each newspaper are far from one another, so no keen competition

# Example 4 (Breakfast Data)

* Using `mdpref` model

```{r}
breakfast<-read.csv("~/MSTAT/STAT8019_Marketing_Analytics/lecture notes/chap4r/b1.csv", header = TRUE,row.names = 2)
head(breakfast[,1:7])
```

## `mdpref` Model

* Set `monotone=T` because it is rank data (ordinal)

```{r}
fit<-cmdpref(breakfast[,2:22],monotone=T,maxit=100)
summary(fit)
```

## Perceptive Map

* Major finding for the MDS is the plot
* Subjectively interpret

```{r}
par(mar=c(2,1,2,1))
plot(fit, xlim=c(-3, 3))
```

* Dim 1 separates the simple-soft vs toasted-hard
* Dim 2 separates the simple-hard vs soft-toasted

4 Major Segments

* Simple
* Hard
* Toasted
* Soft

```{r}
#object plot
plot(fit$score)
text(fit$score,labels=rownames(fit$score),cex=0.8,pos=1)

```

```{r}
#subject plot
plot(fit$corr)
text(fit$corr,labels=rownames(fit$corr),cex=0.8,pos=1)
```

* MDS gives subjective segment and positioning
* MAjor result is the graph to interpret