---
title: "Chapter"
output: html_notebook
---

# Libraries

```{r, message=FALSE}
library(lmtest)
library(car)
library(mlogit)
```

# Example 1 (`book` data)

```{r}

book<-read.csv("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r6/book.csv", header = TRUE)

head(book)
```

Independent Variables:

* `lastpurchase`
* `artbook`

Response Variable:

* `buy`

## Frequency Distribution

```{r}
cbind( Freq=table(book[,"buy"]), percent=prop.table(table(book[,"buy"])),
       Cum.percent=cumsum(prop.table(table(book[,"buy"]))))
```


## Logistic Regression Model

```{r}
logit <- glm(formula=buy ~ artbook + lastpurchase, data = book, family = binomial)
summary(logit)
```

### Likelihood Ratio Test

* Model 0: intercept only
* Model 1: full model

```{r}
#~ 1 intercept only
model0<-glm(buy ~ 1, data = book, family = binomial)
#lrtest() ANOVA table and likelihood ratio test
#lrtest(restricted model, full model)
lrtest(model0,logit)
```

* Model shouldn't be reduced because full model is significant

### Prediction

#### Utility (mu)
```{r}
#predicted utility, xb
predict(logit,book)[1:5]

```

#### Probability


```{r}
#predicted probability
predict(logit,book,type="response")[1:5]
```

```{r}
resp<-cbind(book,
utilty=predict(logit,newdata=book),
p=predict(logit,newdata=book,type="response"))
resp[1001:1002,]
```


## Practical Analysis

* What threshold probability c should be used to define 0 and 1?
* Need to look at the profit driven equation

### Top 50 customers

One option is to look at top 50 customers sorted by `p`

```{r}
pred<-resp[c(-1001,-1002),]
#list the top 50 customers with highest response probabilities
pred[order(pred$p,decreasing=T)[1:50],c("id","p")]
```

### Profit Driven

* Cost of mailing is $1
* Net profit after mailing is $6
* Profit maximizing strategy is to mail all with expected prob > 0

$$6 \times p_c + (-1)\times(1-p_c) > 0$$

leads to $p_c > 1/7$

```{r}
pred$mail<-(pred$p>1/7)
pred1<-pred[pred$mail==T,]
pred1[order(pred1$p,decreasing=T),c("id","p")]
```

```{r}
table(pred$buy)
table(pred$buy,pred$mail)
```

* If using original labels vs profit driven
* Expected net profit of -419 vs 108

# Example 1.c (`chocolate` data)

* Different data structure with 8 profiles
* `n` for number of people presented to
* `c` for choose
* `nc` for not choose


```{r}
load("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r6/chocolate.RData")

chocolate
```


## Pre-processing

* Need to convert into factors
* Check the contrasts

```{r}
chocolate$dark<-factor(chocolate$dark,levels=1:2,labels=c("milk","dark"))
chocolate$soft<-factor(chocolate$soft,levels=1:2,labels=c("chewy","soft"))
chocolate$nuts<-factor(chocolate$nuts,levels=1:2,labels=c("no nuts","nuts"))


contrasts(chocolate$dark)
contrasts(chocolate$soft)
contrasts(chocolate$nuts)
```

## Modelling 

* Need to apply different syntax due to grouped form data

```{r}
mod.form <- "cbind(c,nc) ~ dark+soft+nuts"
logit <- glm(formula=mod.form, family=binomial, data=chocolate)

summary(logit)
```

* Soft/Chewy is the most important factor followed by dark/milk

* The most preferred profile is dark, chewy , and nuts



### Type III ANOVA

```{r}
#type III anova test for each factor
Anova(logit,type="III")
```

* all are significant

## Utlity & Probability Prediction

```{r}

cbind(logit$data,utility=predict(logit,type="link"),p=predict(logit,type="response"))
```


## Test set prediction

```{r}
# scenarios
new<-read.table(text="
dark soft nuts
1 1 2
2 1 1
2 1 2",header=T)
new$dark<-factor(new$dark,levels=1:2,labels=c("milk","dark"))
new$soft<-factor(new$soft,levels=1:2,labels=c("chewy","soft"))
new$nuts<-factor(new$nuts,levels=1:2,labels=c("no nuts","nuts"))

cbind(new,u=predict(logit,newdata=new),p=predict(logit,newdata=new,type="response"))
```

# Example 2 (`laundry` data)

```{r}
laundry <- read.csv("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r6/laundry.csv", header = TRUE)

head(laundry)
```

* `visit` is the unique identifer 
* `alt` is the alternative choices


## Pre-processing

* Convert to factor and relevel for categorical variables

```{r}
laundry$brand<-factor(laundry$brand,levels=1:4,labels=c("Tide","Wisk","Era","Surf"))
laundry$brand<-relevel(laundry$brand,ref="Surf")
contrasts(laundry$brand)
```

## Modelling

* Need to create a `mlogit` object
* `shape=long` data type: the shape of the data.frame: whether long if each row is an alternative or wide if each row is an observation,

### `mlogit` Object Creation

```{r}
#mlogit.dat() create data frame for logit model
#data= input data frame
#choice= choice variable, 0=not, 1=chosen
#shape="long" long shape
#chid.var= choice set ID variable, unique
#alt.var= alternatives ID variable, 1..n
ld <- mlogit.data(data=laundry, choice = "c", 
                  shape = "long",chid.var="visit",
                  alt.var="alt")
```

### Fitting (only `brand`)

* intercept is 0 because for multinomial logistic regression it doesn't matter


```{r}
#mlogit() multinomial logit model
#formula= model, 0 intercept
#data= input data obatined by mlogit.data()
#alt.var= alternatives ID variable
#shape= shape of input data
fit<-mlogit(formula=c~0+brand,data=ld,alt.var="alt",
            shape="long")
summary(fit)
```

#### Test Set & Prediction

* Need to make sure test set has same data structure as train set
* Same pre-processing

```{r}
#new data set
new<-read.table(text="
brand visit alt c
1 999 1 1
2 999 2 1
3 999 3 1
4 999 4 1",header=T)
new$brand<-factor(new$brand,levels=1:4,labels=c("Tide","Wisk","Era","Surf"))
new$brand<-relevel(new$brand,ref="Surf")

contrasts(new$brand)

```


```{r}
#convert new data frame into mlogit.data
newld <- mlogit.data(data=new, choice = "c", 
                  shape = "long",chid.var="visit",
                  alt.var="alt")

#predict() predict choice probabilities
#newdata= new data set
#predict(fit,newdata=newld)

cbind(new,share=predict(fit,newdata=newld))
```

```{r}
#crosstab brand with choice
tb<-table(laundry$brand,laundry$c)
prop.table(tb,margin=1)
```

### Fitting Full Model

```{r}
fit<-mlogit(formula=c~0+brand+price+disp+feat,
            data=ld,alt.var="alt",shape="long")
summary(fit)
```

* Most are significant at a 5% level

#### Coefficients (Part-worths)

```{r}
#estimated coefficients
fit$coefficients
```

* Since price is continuous, the part worth depends on the price

$$PW_{price}=-0.51 \times price$$

```{r}
#calcualte part-worth for each price
pu<-cbind(price=unique(ld$price),part.worth=unique(ld$price)*fit$coefficients["price"])
pu[order(pu[,"price"]),][1:10, ]
```

#### Test Set Predictions

```{r}
# scenarios
new<-read.table(text="
brand disp feat price visit alt c
1 1 0 5 999 1 1
1 0 0 4.5 999 2 1
3 1 1 6 999 3 1
2 0 1 4 999 4 1
",header=T)
new$brand<-factor(new$brand,levels=1:4,labels=c("Tide","Wisk","Era","Surf"))
new$brand<-relevel(new$brand,ref="Surf")

newld <- mlogit.data(data=new, choice = "c", 
                     shape = "long",chid.var="visit",
                     alt.var="alt")

cbind(new,share=predict(fit,newdata=newld))
```

# Example 3 (data mining challenge)

* Computing **uplift**
* Two options, use 2 models or 1 model


```{r}
mydata <- read.csv("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r6/dm.csv")

head(mydata)
```

## Pre-processing

```{r}
#remove Womens E-mail
mydata<-mydata[mydata$segment!="Womens E-Mail",]

#reset the factor levels
mydata$conversion<-factor(mydata$conversion)
mydata$segment<-relevel(mydata$segment,ref="No E-Mail")

mydata$segment<-factor(mydata$segment)
table(mydata$segment)

contrasts(mydata$segment)
contrasts(mydata$conversion)
```

```{r}
table(mydata$conversion)
```


## Separate Models (2 models)

```{r}
#separated models
model<-conversion ~ newbie+recency+history_segment+
                    channel+history+zip_code+mens+womens

#logistic model applied to treatment group
fit.trt<-glm(model,data=mydata[mydata$segment=="Mens E-Mail",],family=binomial)
summary(fit.trt)

```

```{r}
#logistic model applied to control group
fit.ctl<-glm(model,data=mydata[mydata$segment!="Mens E-Mail",],family=binomial)
summary(fit.ctl)
```

### Prediction and Uplift

```{r}
#predict the response probability of all respondents when treatment is applied
p.trt<-predict(fit.trt,newdata=mydata,type="response")
#predict the response probability of all respondents when treatment is not applied
p.ctl<-predict(fit.ctl,newdata=mydata,type="response")
#uplift = treatment probability - control probability
uplift<-cbind(mydata,p.trt,p.ctl,uplift=p.trt-p.ctl)
head(uplift[order(-uplift$uplift),c("p.trt","p.ctl","uplift"),])
```

```{r}
# for customer 32653
mydata[rownames(mydata)==32653,]
```


## Combined model

* Cross interact the segment with each of the variable

```{r}
#combined model
model<-conversion ~ newbie*segment+recency*segment+history_segment*segment+
                    channel*segment+history*segment+zip_code*segment+
                    mens*segment+womens*segment

fit.comb<-glm(model,data=mydata,family=binomial)
summary(fit.comb)
```


### Prediction and Uplift

```{r}
pdata<-mydata
#predict the response probability of all respondents when treatment is applied
pdata$segment<-"Mens E-Mail"
p.trt<-predict(fit.comb,newdata=pdata,type="response")
#predict the response probability of all respondents when treatment is not applied
pdata$segment<-"No E-Mail"
p.ctl<-predict(fit.comb,newdata=pdata,type="response")
uplift<-cbind(mydata,p.trt,p.ctl,uplift=p.trt-p.ctl)
head(uplift[order(-uplift$uplift),c("p.trt","p.ctl","uplift"),])
```


