---
title: "Chapter 7"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r}
library(forecast)
library(plyr)
library(ggplot2)
library(patchwork)
```

# Example 3 (Minivan)

```{r}
automob<-read.csv('/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r7/automob.csv',header=TRUE)
#define model as factor, instead of numeric values
#automob$model<-factor(automob$i,labels=c("Caravan","Voyager","APV","Silhouette"))
head(automob)
```

## Pre-processing

* Using APV as example

```{r}
caravan<-automob[automob$i==1,]
voyager<-automob[automob$i==2,]
apv<-automob[automob$i==7,]
silhouette<-automob[automob$i==8,]
```

* Normalize data based on a record 

```{r}
#normalized data
apv$nyt<-apv$yt/60772
apv$nt<-apv$t/6
```



## EDA
```{r}
#plot of sales and cumulative sales
#Caravan    Voyager    APV        Silhouette
mxy<-max(apv$y)
mxyt<-max(apv$yt)
 
ggplot(data=apv, aes(x=t)) +
       geom_line(aes(y=y,color="y")) +
       geom_point(aes(y=y,color="y"))+
       geom_line(aes(y=yt/mxyt*mxy,color="yt")) +
       geom_point(aes(y=yt/mxyt*mxy,color="yt"))+
       scale_y_continuous(sec.axis = sec_axis(~.*mxyt/mxy,name="yt")) +
       scale_colour_manual(values = c("blue", "red")) +
       theme(legend.position="bottom",legend.title=element_blank()) +
       ggtitle("APV")
```


## Modelling

### ADBUGD Model

* Uses cumulative data from previous entry
* S-shape

```{r}
#ADBUG model
fit <- nls(f=nyt~a+b*nt^c/(nt^c+d),data=apv,
                start=list(a=0,b=1.89,c=0.89,d=3))
summary(fit)

#r2
r2<-1-(sum(residuals(fit)^2)/sum((fit$m$lhs()-mean(fit$m$lhs()))^2))
r2
```

* Almost perfect fit from model onto the data

```{r}
#prediction
nd<-read.table(text="t
11
12
13
14
15
",header=T)

nd$nt<-nd$t/6

pred<-cbind(apv,predict=predict(fit)*60772)

ggplot(data=pred, aes(x=t)) +
  geom_point(aes(y=yt)) + 
  geom_line(aes(y=predict)) + ggtitle("ADBUDG Model Fitting on Cumulative Previous Data")
```


## Diffusion Model - Bass Model

```{r}
#difussion model
#Caravan    Voyager    APV        Silhouette
#Bass model
fit1 <- nls(f=y~a*n+(b-a)*yt-b/n*(yt^2),
            start=list(a=0,b=0.5,n=10000),
            data=apv)
summary(fit1)

#r2  
r2<-1-(sum(residuals(fit1)^2)/sum((fit1$m$lhs()-mean(fit1$m$lhs()))^2))
r2
```

```{r}
ggplot(data=apv, aes(x=t)) +
  geom_point(aes(y=y))+ 
  geom_line(aes(y=predict(fit1))) +
  ggtitle("Bass Model on Sales Data")
```


# Example 4 (Pulses)

```{r}
pakinstan<-read.table('/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r7/pakinstan.dat',header=TRUE)

head(pakinstan)
```

## EDA

```{r}
#time series plot
p1 <- ggplot(data=pakinstan,aes(x=year)) +
  geom_line(aes(y=gram)) + 
  geom_point(aes(y=gram))

p2 <- ggplot(data=pakinstan,aes(x=year,y=mung)) +
  geom_line() + 
  geom_point()

p1 + p2
```

## Modelling

### Exponential Smoothing

```{r}
#holt() exponential smoothing forecasts
#y= input time series
#h= no of ahead forecasts
#init= initial state(s) estimation method
#   "simple" initial states is not optimized
#beta=FALSE, simple exponential smoothing
fit<-holt(y=pakinstan$gram, h=6, initial="simple",
          beta=FALSE)
summary(fit)

```

```{r}
#fit$model$SSE, SSE
#fit$x, input time series
r2<-1-fit$model$SSE/sum((fit$x-mean(fit$x))^2)
r2
fit$model$states
```

```{r}
plot(fit)
```


* alpha close to 1, implies current values contributes a lot to predict future
* Lt depends heavily on Yt

### Holt's Linear Exponential Smoothing

```{r}
#Holt's linear exponential smoothing
#beta=NULL (by default), linear exponential smoothing
fit<-holt(y=pakinstan$gram, h=6, initial="simple",beta=NULL)
summary(fit)

```


```{r}
plot(fit)
r2<-1-fit$model$SSE/sum((fit$x-mean(fit$x))^2)
r2
fit$model$states
```

* Beta is close to 0, implying a smooth trend


### Exponential Smoothing (Mung)

```{r}
#simple exponential smoothing
fit<-holt(y=pakinstan$mung, h=6, initial="simple",beta=FALSE)
summary(fit)
r2<-1-fit$model$SSE/sum((fit$x-mean(fit$x))^2)
r2
fit$model$states

```

```{r}
plot(fit)
```


### Holt's Linear Exponential Smoothing (Mung)


```{r}

#Holt's linear exponential smoothing
fit<-holt(pakinstan$mung, h=6, initial="simple",beta=NULL)
summary(fit)
r2<-1-fit$model$SSE/sum((fit$x-mean(fit$x))^2)
r2
plot(fit)
fit$model$states
```

* 6 years forecasted price tend to rise


# Example 5 (Load Forecast)

```{r}
power<-read.csv('/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r7/power.csv',header=TRUE)
#convert date into a date value
power$date<-as.Date(power$date,"%d%b%Y")

head(power)
```


## EDA

```{r}
#series plot
ggplot(data=power,aes(x=date,y=demand))+
  geom_line()+
  geom_point()
```

## Pre-processing

* `hw()` function only takes in a time series `ts` object
* Need to convert it

```{r}
#ts() create a time series
#frequency= no. of observations in a season/period
#start= starting year and month
ts.power<-ts(data=power[1:168,"demand"],frequency=12,start=c(2002,1))

ts.power
```


## Modelling

### Holt-Winters Additive 

```{r}
#Holt-Winters additive model
#hw() Holt-Winters linear exponential smoothing with seasonal effects
#seasonal="add", additive model
fit<-hw(y=ts.power,seasonal="add",h=24,initial="simple")
summary(fit)
r2<-1-fit$model$SSE/sum((fit$x-mean(fit$x))^2)
r2
fit$model$states[169,]


```

```{r}
plot(fit)

forecast(fit)

# point forecast
forecast(fit)$mean
```

### Holt-Winters Multiplicative Model

```{r}
#Holt-Winters multiplicative model
#seasonal="mul", multiplicative model
fit<-hw(ts.power,seasonal="mul",h=24,initial="simple")
summary(fit)
r2<-1-fit$model$SSE/sum((fit$x-mean(fit$x))^2)
r2
round(fit$model$states[169,],digits=2)
```

```{r}
plot(fit)

forecast(fit)
forecast(fit)$mean
```


# Example 10 (Marketing Mix Modelling)

* Find the optimal salespeople allocation to maximize profit
* If total salespeople is restricted to 80, find the optimal allocation

```{r}
#load R data sets
load(file="/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r7/eg10.RData")
head(product)

```

## Normalization

```{r}
#normalization
product$nsalesforcea<-product$salesforcea/24
product$na<-product$salesa/37200
product$nsalesforceb<-product$salesforceb/27
product$nb<-product$salesb/38200
```


## EDA

```{r}
ggplot(data=product)+
  geom_line(aes(x=nsalesforcea,y=na,color="A"))+
  geom_point(aes(x=nsalesforcea,y=na),color="red")+
  geom_line(aes(x=nsalesforceb,y=nb,color="B"))+
  geom_point(aes(x=nsalesforceb,y=nb),color="blue") +
  labs(x="norm. salesforce",y="norm. sales") +
  scale_color_manual(labels = c("A", "B"), values = c("red", "blue")) +
  theme(legend.position = "right")+
  theme(legend.title = element_blank())
```


## Modelling

* Fit 2 separate ADBUDG model for product A and B

### Product A
```{r}
#product A
fita <- nls(f=na~a+b*nsalesforcea^c/(nsalesforcea^c+d),data=product,
            start=list(a=0.4,b=1,c=2.6,d=0.9))
summary(fita)
r2<-1-sum(residuals(fita)^2)/sum((fita$m$lhs()-mean(fita$m$lhs()))^2)
r2
```

### Product B

```{r}
#product B
fitb <- nls(f=nb~a+b*nsalesforceb^c/(nsalesforceb^c+d),data=product,
            start=list(a=0.1,b=1,c=2.9,d=0.4))
summary(fitb)
r2<-1-sum(residuals(fitb)^2)/sum((fitb$m$lhs()-mean(fitb$m$lhs()))^2)
r2
```

```{r}
ggplot(data=product)+
  geom_line(aes(x=nsalesforcea,y=predict(fita),color="A"))+
  geom_point(aes(x=nsalesforcea,y=na),color="red")+
  geom_line(aes(x=nsalesforceb,y=predict(fitb),color="B"))+
  geom_point(aes(x=nsalesforceb,y=nb),color="blue") +
  labs(x="norm. salesforce",y="pred. norm. sales") +
  scale_color_manual(labels = c("A", "B"), values = c("red", "blue")) +
  theme(legend.position = "right")+
  theme(legend.title = element_blank())

```

## Optimization

### Without Constraint 

```{r}
#objective function
#multiple parameters
#nsf is a vector
#nsf[1] is the first parameter
#nsf[2] is the second parameter
#...
net<-function(nsf) {
  nd<-data.frame(nsalesforcea=nsf[1])
  nsalesa<-predict(fita,nd)
  margina<-0.72
  salesa<-nsalesa*37200
  # reverse normalization
  sfa<-nsf[1]*24
  costa<-sfa*63
  grossa<-margina*salesa
  neta<-grossa-costa

  nd<-data.frame(nsalesforceb=nsf[2])
  nsalesb<-predict(fitb,nd)
  marginb<-0.62
  salesb<-nsalesb*38200
  # reverse normalization
  sfb<-nsf[2]*27
  costb<-sfb*63
  grossb<-marginb*salesb
  netb<-grossb-costb

  net=neta+netb
  
#print(cbind(nsalesa,salesa,nsalesb,salesb,sfa,sfb,grossa,grossb,neta,netb,net))
  return(net)
}
net(c(3,2))
#optim() General-purpose Optimization
#        optimization with multiple parameters
#fn= objective function
#par= initial estimates
#control=list(fnscale=-1)  maximization
op<-optim(par=c(3,2),fn=net,control=list(fnscale=-1))
op

net(op$par)
```

### With Constraint


```{r}
#constrOptim() linearly constrained optimization
#theta= initial estimates
#f= objective function
#ui= constrain matrix
#ci= constrain vector
#    ui*beta-ci >= 0
#    80>=24*nsfa+27*nsfb
#    <=>
#    -24*nsfa-27*nsfb+80>=0
#grad=NULL gradient function not provided
op<-constrOptim(theta=c(1,1),f=net,
                control=list(fnscale=-1),
                ui=c(-24,-27),ci=-80,grad=NULL)
op

net(op$par)
```

### Plot

```{r}
net<-function(nsf) {
  nd<-data.frame(nsalesforcea=nsf[1])
  nsalesa<-predict(fita,nd)
  margina<-0.72
  salesa<-nsalesa*37200
  sfa<-nsf[1]*24
  costa<-sfa*63
  grossa<-margina*salesa
  neta<-grossa-costa
  
  nd<-data.frame(nsalesforceb=nsf[2])
  nsalesb<-predict(fitb,nd)
  marginb<-0.62
  salesb<-nsalesb*38200
  sfb<-nsf[2]*27
  costb<-sfb*63
  grossb<-marginb*salesb
  netb<-grossb-costb
  
  net=neta+netb
  
  #print(cbind(nsalesa,salesa,nsalesb,salesb,sfa,sfb,grossa,grossb,neta,netb,net))
  return(net)
}


profile<-data.frame(matrix(rep(0,3*7*7),byrow=T,nrow=7*7))
colnames(profile)<-c("nsalesforcea","nsalesforceb","net")
for (i in 0:6) {
  for (j in 0:6) {
    profile[i*7+j+1,1:2]<-cbind(i/2,j/2)
    profile[i*7+j+1,3]<-net(profile[i*7+j+1,1:2])
  }
}
profile$nsalesforcea<-factor(profile$nsalesforcea)
ggplot(data=profile)+
  geom_line(aes(x=nsalesforceb,y=net,group=nsalesforcea,colour=nsalesforcea))+
  labs(x="norm. salesforce B",y="net profit",colour="norm. sf A") +
    theme(legend.position="right" ) 

```


