---
title: "Chapter 3"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(magrittr)
library(dplyr)
library(tidyverse)
library(psych)
library(reshape2)
library(GPArotation)
```


# Brand Data

There are 9 variables here. Can the data be described in less number of variables?

## EDA

```{r}
load("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/chap3r/brand.RData")

head(brand.ratings)
```

```{r}
describe(brand.ratings, skew = F, ranges = F)
```

```{r}
brand.ratings %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density()  
```

### Correlation

```{r}
print(cor(x=brand.ratings[, 1:9]), digits = 2)
```

* Perform, serious, and leader are positively correlated
* Trendy and latest are positively correlated
* Bargain, value, and rebuy are positively correlated

## Factor Analysis using PCA method (THIS IS NOT PCA but PCFA)

* `principal` uses standardized values (using the correlation matrix), hence the variance of each variable is 1
* Since `n=9`, the total variance is 9

```{r}
fit <- principal(r=brand.ratings[, 1:9], nfactors = 9, rotate = "none")
fit
```

* First 4 factors already explain 76% of the variance
* If we use **Latent Root Criteria**, then only choose factors that are > 1, which is this case are the first 3 factors
* If scree plot, elbow at 3, so first 2 factors

### Scree Plot

```{r}
scree(rx=brand.ratings[, 1:9], factors = F)
```

### Decide on 3 Factors

* Based on all the analysis, a 3 factor model is decided to be used

```{r}
# refit the 3 factor model

fit <- principal(r=brand.ratings[, 1:9], nfactors = 3, rotate = "none")
```

```{r}
fit$loadings
```


* This implies the following:
  * $Z_{perform} = 0.41F_{1} + 0.61F_{2} - 0.04F_{3} + \epsilon_1$
  * $Z_{leader} = 0.36F_{1} + 0.76F_{2} + 0.10F_{3} + \epsilon_2$
  * ...


```{r}
fit$communality
```

```{r}
fit$uniquenesses
```


* `h2` is the communality and `u2` is the uniqueness, sum is 1
  * 54% of variance of `perform` explained by 3 factors, $(0.41)^2+(0.61)^2+(-0.04)^2 = 0.54$
  * ...

### Apply Rotation with `rotate=varimax`

* Rotation doesn't change the communalities and uniqueness
* Total variance explain will also still be the same
* Loadings will be different

```{r}
#rotate="varimax" varimax rotation
#method="regression" factor scores estimated by regression method
fitr <- principal(r=brand.ratings[,1:9],nfactors=3,rotate="varimax",method="regression")
fitr
```

In terms of the rotated loadings:

* $Z_{perform} = 0.72F^*_1 + 0.10F^*_2 - 0.09F^*_3 + \epsilon_1$
* ...

We may interpret the rotated loadings as follows:

* Factor 1
  * Loads heavy on (descending order) `leader, performs, serious`
  * Loads negative on `fun`
  * Leader factor
  * Customers look at leader, then high score for perform, leader, serious, low score for fun

* Factor 2
  * Loads heavy on `bargain, value, rebuy`
  * Value factor

* Factor 3
  * Loads heavy on `latest, trendy`
  * Latest factor

#### Rotation Matrix

* This is the rotation matrix **T** that is in $LT=L^*$

```{r}
fitr$rot.mat[c(1,2,3), c(2,1,3)]
```

### Weights and Scores

```{r}
fitr$weights
```

```{r}
head(scale(brand.ratings[, 1:9]))
```


* $Z_{perform}$ is the standardized value of Z

* Observation 1
  * Factor 1 score $f^*_1 = 0.32Z_{perform} + ... +0.11Z_{rebuy}$
  * $f^*_1 = 0.32(-0.78) + ... +0.11(0.89) = -0.86$

```{r}
head(fitr$scores)
```

### Mean Scores by Brand

```{r}
scores<-cbind(brand.ratings["brand"],fitr$scores)

#group means of factor scores
tb<-aggregate(.~brand,data=scores,FUN=mean)
tb
```

```{r}
#plot of the group means
tbm<-melt(tb,id.vars='brand')

ggplot(tbm, 
       aes(x = variable, y = value, group = brand, colour = brand)) + 
  geom_line(aes(linetype=brand))+
  geom_point(aes(shape=brand)) +
  geom_hline(yintercept=0) +
  labs(x=NULL,y="mean") +
  scale_shape_manual(values=1:nlevels(tbm$brand)) +
  scale_color_manual(values=c(1,2,2,3,4,5,5,3,3,1))
```

**Interpretation**

* Brand F, G, high at Value (factor 2)
* Brands B, C high at Leader (factor 1)
* Brands E: high at Latest (factor 3) and high at Value (factor 2)
* Brands A, J: low at Leader (factor 1) ⇒ high at Fun
* Brands D, H, I: high at Latest (factor 3) and low at Value (factor 2)


### 2-Factor Solution

```{r}
#2-factor solution
fit <- principal(r=brand.ratings[1:9], nfactors=2,rotate="varimax")

#plot of factor loadings
ld<-data.frame(fit$loadings[,1:2])
ggplot(data=ld,aes(x=RC1,y=RC2))+
  geom_point()+
  geom_text(aes(label=rownames(ld),vjust=1))+
  geom_vline(aes(xintercept=0))+
  geom_hline(aes(yintercept=0))+
  coord_cartesian(xlim = c(-1,1),ylim=c(-1,1)) 
```

* Consistent with results from the 3 factor solution, similar grouping
* Even with 2 dimensions, structure still retain
* Leader, serious, perform are correlated
* Value, rebuy and bargain are correlated
* Latest, trendy are correlated
* Fun on its own

```{r}
#means factor scores
sc<-data.frame(fit$scores,brand=brand.ratings$brand)
tb<-aggregate(.~brand,data=sc,FUN=mean)
#plot
ggplot(data=tb,aes(x=RC1,y=RC2))+
  geom_point()+
  geom_text(aes(label=brand,vjust=1))+
  geom_vline(aes(xintercept=0))+
  geom_hline(aes(yintercept=0))
```

* Comparing the factor loadings plot and the factor scores plot
  * B, C
* Leader, serious, perform
  * F, G
* Value, rebuy, bargain
  * A, J
* Fun
  * D, H, I
* Latest, trendy
  * E
* No strong characteristic

#### Clustering using Ward's Method

Based on plot above, it seems like there are 5 groups. Clustering can be done to see if it is consistent with above groupings.

```{r}
dist<-dist(sc[1:2],method="euclidean")^2
fitc <- hclust(dist, method="ward.D")
sol <- cutree(fitc, k=5)
sol<-data.frame(sol,sc)
sol$sol<-factor(sol$sol)
ggplot(data=sol,aes(x=RC1,y=RC2,colour=sol))+
  geom_point(aes(shape=sol)) +
  geom_vline(aes(xintercept=0))+
  geom_hline(aes(yintercept=0))
```

```{r}
table(sol[c("sol", "brand")])
```

# Preference Data

Instead of Principle Component method, the **Maximum Likelihood** method will be used for obtaining the Factors.

## EDA

```{r}
pref<-read.csv("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/chap3r/pref.csv", header = TRUE)

colnames(pref) <- c("id", "taste", "goodbuy", "flavor", "suitable", "energy")

head(pref)
```

```{r}
describe(pref, skew = F, ranges = F)
```

```{r}
pref[, 2:6] %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density()  
```


```{r}
as.data.frame(round(cor(pref[, 2:6]), 2))
```

* Taste and Flavor (correlation = 0.899), Good buy for money and Provides lots of energy (correlation = 0.728) form groups

* Suitable for snack closer to (Good buy for money and Provides lots of energy) (correlation = 0.730 and 0.712) than (Taste and Flavor) (correlation = 0.623 and 0.584)

* Linear relationships between the variables can be explained in terms of 2 (or 3) common factors

### Scree Plot

```{r}
#scree plot
scp<-scree(pref[2:6],factors = F)
```

These are the eigen values estimated by the principle component method, unrotated.

```{r}
scp$pc
```

## Factor Analysis using MLE

Unrotated
```{r}
fit <- fa(r=pref[, 2:6], nfactors = 2, rotate = "none", fm = "ml")
fit
```

```{r}
fit$communalities
```

* Based on the Chi-Squared test above, 2 factors are sufficient, 83% of variance can be explained
* R always shows the Barlett Corrected $\chi^2$ statistic

### Factor Loadings

```{r}
#factor loadings
ld<-data.frame(fit$loadings[,1:fit$factors])
ggplot(data=ld,aes(x=ML1,y=ML2))+
  geom_point()+
  geom_text(aes(label=rownames(ld),vjust=1))+
  geom_vline(aes(xintercept=0))+
  geom_hline(aes(yintercept=0))+
  coord_cartesian(xlim = c(-1,1),ylim=c(-1,1)) 
```

### With Rotation

```{r}
#varimax rotation
#scores="regression" regression method for factor scores
fitr<-fa(r=pref[2:6],nfactors=2,rotate="varimax",fm="ml",scores="regression")
print(fitr, digits = 3)
```

* Factor 1
  * Good buy for money (X2), Suitable for snack (X4) and Provides lots of energy (X5) define factor 1 (high loadings on factor 1, small or negligible loadings on factor 2)
  * Nutritional factor
  
* Factor 2
  * Taste (X1) and Flavor (X3) define factor 2 (high loadings on factor 2, small or negligible loadings on factor 1)
  * Taste factor

```{r}
#loading plot
ld<-data.frame(fitr$loadings[,1:fitr$factors])
ggplot(data=ld,aes(x=ML2,y=ML1))+
  geom_point()+
  geom_text(aes(label=rownames(ld),vjust=1))+
  geom_vline(aes(xintercept=0))+
  geom_hline(aes(yintercept=0))+
  coord_cartesian(xlim = c(-1,1),ylim=c(-1,1)) 
```


```{r}
fitr$weights
```

```{r}
head(fitr$scores)
```

* As usual, communalities and total variance explained doesn't change.

### Perceptual Plot

* Individual plots of each observation with Factor Scores

```{r}
#factor scores
sc<-data.frame(fitr$scores)
ggplot(data=sc,aes(x=ML2,y=ML1))+
  geom_point()+
  geom_vline(aes(xintercept=0))+
  geom_hline(aes(yintercept=0))+
  geom_text(aes(label=pref$id,vjust=2))+
  coord_cartesian(ylim=c(-3,2)) 

```

** ML2 = nutrition , ML1 = taste **

* Most people preferred good at both nutrition and taste
* There are extreme cases
  * Case 44 and 4
* Prefer very good at taste even if very bad at nutrition
  * Case 22 and 16
* Prefer very good at nutrition even if very bad at taste
  * No people preferred product that is bad at both nutrition and taste

