---
title: "Chapter 5"
output: html_notebook
---

# Libraries

```{r}
library(conjanal)
library(ggplot2)
library(reshape2)
```

# Example 1: `cleanser` data

* 18 for most preferred  (because have 18 profiles)

* 1 for least preferred

* Highest values must be most preferred and lowest must be least preferred for the program to work

* Each row is a profile. Last 3 rows for prediction (choice simulation only)

* Must convert to factor for program to work

* Sum of part-worth is 0 (set by using `contrasts`)

```{r}
clean <-read.csv("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r5/cleanser.csv")

head(clean)
```

```{r}
#define factors
clean$color<-factor(clean$color,levels=c(1,2,3),labels=c("red","blue","yellow"))
clean$form<-factor(clean$form,levels=c(1,2,3),labels=c("cubic","cylinder","spherical"))
clean$scent<-factor(clean$scent,levels=c(1,2),labels=c("yes","no"))
```

```{r}
head(clean)
```


## Data Preparation

* Next chapter is choice based

* Monotonic Anova

### Step 1: Set Contrasts for Sum of Part Worths to be 0

```{r}
#set contrasts, sum=0
contrasts(clean$color)<-contr.sum(3)
contrasts(clean$form)<-contr.sum(3)
contrasts(clean$scent)<-contr.sum(2)
```

### Step 2: Train, Test Split 
```{r}
# train test split
# scenarios
csimp<-clean[19:21,1:3]
clean<-clean[1:18,]
```

```{r}
clean
```


```{r}
csimp
```

### Step 3: Create Response Matrix and Profile of Training Data

* The respondents column of the training data

```{r}
#define response matrix. Need to convert as matrix
resp<-as.matrix(clean[,4:5])

#define profiles
profile<-clean[,1:3]
```


### Step 4: Apply Conjoint Analysis

* Because using ranking, `monotone=True`. Monotonic transformation.

```{r}
#conjanal() metric conjoint analysis
#resp= matrix of responses
#profile= data frame of designs
#iv= factors
fit.conj<-conjanal(resp=resp,profile=profile,monotone = T)
summary(fit.conj$lm)
```

```{r}
#contrasts of factors
contrasts(clean$form)
contrasts(clean$color)
contrasts(clean$scent)
```

```{r}
#part-worth
fit.conj$part
#relative importance
fit.conj$imp
```

For `r1`:

* Form is the most important factor accounting for 51% 
* Second most important is color
* Yellow is the most important color as it has the highest part worths among all colours
* Cylinder most preferred
* No scent
* Optimal profile is yellow, cylinder, no scent

For `r2`:

* Color is the most important factor accounting for 42%
* Red, cubic, no scent is preferred

## Predict Utility

* Predict for individual customer

```{r}
csimp
```

### Train Test set predict

```{r, warning=FALSE}
#predicted utilities
cbind(profile,pred=predict(fit.conj$lm,newdata=profile))

cbind(csimp,pred=predict(fit.conj$lm,newdata=csimp))
```

* negative predictions do happen because of linear model, no bounds

## Market Share (choice simulation)

### Maximum Util

```{r}
#maxconj() maximum utility
#conjanal= result by conjanal()
#simp= scenarios data frame
maxconj(conjanal=fit.conj,simp=csimp)
```

### BTL

```{r}
#btlconj() BTL model
btlconj(conjanal=fit.conj,simp=csimp)
```

### Logit

```{r}
#logitconj() logit model
logitconj(conjanal=fit.conj,simp=csimp)
```

# Example 2 (sneakers data)

* Full profile approach
* Fractional design (9 profiles only instead of 27)

```{r}
sneaker <- read.csv("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/lecture notes/r5/sneaker.csv")

sneaker
```

* Rows describe the profile

* Columns describe the responses

## Data Prep

```{r}
#define factors
sneaker$COL1<-factor(sneaker$COL1,levels=c(1,2,3),labels=c("rubber","polyurethane","plastic"))
sneaker$COL2<-factor(sneaker$COL2,levels=c(1,2,3),labels=c("leather","canvas","nylon"))
sneaker$COL3<-factor(sneaker$COL3,levels=c(1,2,3),labels=c("30","60","90"))

#set contrasts, sum=0
contrasts(sneaker$COL1)<-contr.sum(3)
contrasts(sneaker$COL2)<-contr.sum(3)
contrasts(sneaker$COL3)<-contr.sum(3)

#define response matrix
resp<-as.matrix(sneaker[,4:63])

#define profiles
profile<-sneaker[,1:3]
```

```{r}
sneaker
```

## ANOVA Model

* Has numerical and categorical, so ANOVA, not MONOANOVA

```{r}
#ANOVA model for first respondent
fit.lm<-lm(COL4 ~ COL1 + COL2 + COL3, data=sneaker)
summary(fit.lm)
```

* Not ranked data, so `monotone=F`

```{r}
fit.conj<-conjanal(resp,profile,monotone=F)
fit.summary<-summary(fit.conj$lm)
fit.summary$`Response COL4`
```

### Part Worth and VarImp

```{r}
#display the first respondent, change 1 to 2 for 2nd respondent
fit.conj$part[,1]
fit.conj$imp[,1]
```

## Cluster Analysis

* Cluster the customers based on variable importance
* transpose data structure and then sscale it


### Hclust

```{r}
# standardize relative importance
imp<-data.frame(scale(t(fit.conj$imp)))
#Ward's method
dist<-dist(imp,method="euclidean")^2
clust <- hclust(dist, method="ward.D")

ggplot(mapping=aes(x=1:length(clust$height),y=clust$height))+
  geom_line()+
  geom_point()+
  labs(x="stage",y="height")

```

* 3 clusters recommended

```{r}
#par(mar=c(1,4,1,1))
plot(clust,labels=F,hang=-1,sub="",xlab="",main="")
```

### K-means

* Optimized grouping

```{r}
set.seed(123)
fit1<-kmeans(x=imp,centers=3)
fit1
tb<-fit1$centers

tb<-data.frame(cbind(tb,cluster=1:3))
tbm<-melt(tb,id.vars='cluster')
tbm$cluster<-factor(tbm$cluster)
ggplot(tbm, 
       aes(x = variable, y = value, group = cluster, colour = cluster)) + 
  geom_line(aes(linetype=cluster))+
  geom_point(aes(shape=cluster)) +
  geom_hline(yintercept=0) +
  labs(x=NULL,y="mean")
```

**Refer to lecture notes:**

* Cluster 1: Price is most important

* Cluster 2: Upper is most important

* Cluster 3: Sole is most important.