---
title: "Tutorial 4"
output: 
  html_notebook:
    toc: true
---

# Libraries

```{r, message=FALSE}
library(dplyr)
library(smacof)
library(cmdpref)
library(ggplot2)
```


# Question 2

* Small number is similar, large is different
* Can straight away use this dataset to `mds()` function

```{r}
load(file='/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/tutorial notes/t4/TUT04.RData')
```

```{r}
head(b2b)
```

## Part (a)

Available scales are `"ratio", "interval", "ordinal", "mspline".`

* Dataset is ranked, so use "ordinal" (nonmetric MDS)

## Part (b)

```{r}
fit1 <- mds(b2b, ndim = 1, type = "ordinal")

fit1

fit2 <- mds(b2b, ndim = 2, type = "ordinal")

fit2 

fit3 <- mds(b2b, ndim = 3, type = "ordinal")

fit3
```

* 2-dimensions have the lowest stress

## Part (c)

Perceptual map

```{r}
plot(fit2, cex=2)
```


## Part (d)

* Differences of asian and western customers
* Dimension 2 separates the **asian and western countries**
* Dimension 1 separates the **northern and southern hemisphere**, culture of the countries

```{r}
# coordinates
fit2$conf
```

## Cluster Analysis

* `b2b` can also be treated as a distance matrix
* apply `hclust` to it

```{r}
fit <- hclust(as.dist(b2b), method = 'ward.D')
plot(fit)
```

* can first use mds to visualize map
* then use hclust 
* hclust only give grouping but mds give direction and visualize in map


# Question 3

* MDS unfolding model means *Multidimensional Scaling Preference (MDPREF)* model

```{r}
crisps<-read.csv("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/tutorial notes/t4/TUT07.csv", header = TRUE,row.names = 1)
head(crisps)
```

## Part (a)

* Using metric MDS

```{r}
crisps <- t(crisps)
```

```{r}
fit5 <- cmdpref(crisps, ndim = 9)
summary(fit5)
```

* Using 2 dimensions explains 42.9% of variation
* Using 5 dimensions explains 72% of variation

## Part (b)

* Using non metric MDS, set `monotone=T`
* Ordinal scale, rank data
* 2 dimensions assumed

### (i)

```{r}
fit2 <- cmdpref(crisps, monotone = T)
summary(fit2)
```

* Monotine using 2 dim explains 64% of variance

### (ii)

```{r}
fit2$score
```

```{r}
head(fit2$corr)
```

### (iii)

```{r}
plot(fit2, xlim=c(-2,3))
```

* No clear goruping of consumers but there are for crisps
* Crisp 2 and 6 are competing
* Crisp 4 and 3 are competing
* Others are clustered

```{r}
plot(fit2$score)
```

