---
title: "Tutorial 6"
output: html_notebook
---

# Libraries

```{r, message=FALSE}
library(mlogit)
library(lmtest)
```

# Question 1

```{r}
bank<-read.csv("/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/tutorial notes/t6data/bank.csv", header = TRUE)
head(bank)
```

## Part (a)

```{r}
# factor and relevel

bank$switch<-factor(bank$switch,labels=c("no","yes"))
bank$switch<-relevel(bank$switch,ref="no")
contrasts(bank$switch)
```

```{r}
logit <- glm(formula=switch ~ 
               appvtime+branch+odprice+guanxi+sow,
               data = bank, family = "binomial")
summary(logit)
```


```{r}
logit$coefficients
```

## Part (b)

H1. Perceived service quality (assessed by the time required to process applications) is associated
with a business customer’s intention to stay with a service provider. -> True and significant

H2. The higher the number of bank branches, the less likely that a customer will intend to switch to
other service providers. -> Not significant

H3. For a given level of service quality, the product price differential relative to competitors is
negatively associated with customer loyalty. -> True and significant

H4. A customer’s intention to switch from a service provider is negatively associated with their
perception of guanxi with that provider. -> True and significant

H5. The higher the extent of the existing SOW, the less likely a customer will intend to switch
service providers. -> Not significant

## Part (c) RELATIVE IMPORTANCE

* Because they are continous variable can just use the abs minus the range of the possible levels

```{r}
#c
#appvtime, 1-3
range1<-abs(logit$coefficients[2]*2)
#branch, 0-2
range2<-abs(logit$coefficients[3]*2)
#odprice, 0.5-1.5
range3<-abs(logit$coefficients[4])
#guanxi, 1-7
range4<-abs(logit$coefficients[5]*6)
#sow, 1-4
range5<-abs(logit$coefficients[6]*3)

range<-c(range1,range2,range3,range4,range5)
#range<-abs(logit$coefficients[-1]*c(2,2,1,6,3))
range/sum(range)
```


## Part (d)

```{r}
nd<-data.frame(guanxi=4,sow=1,odprice=0.5,appvtime=3,branch=0)
predict(logit,newdata=nd)
predict(logit,newdata=nd,type="response")
```


# Question 2

```{r}
load(file="/Users/jasonchan/MSTAT/STAT8019_Marketing_Analytics/tutorial notes/t6data/abb.RData")
```


## Part (a)

```{r}
#a
ld <- mlogit.data(data=abb, choice = "choice", 
                  shape = "long",chid.var="ID",
                  alt.var="altid")

fit<-mlogit(formula=choice~0+company,data=ld,alt.var="altid",
            shape="long")
summary(fit)
```

### Prep Test set and make prediction

* make sure the contrasts are the same as that original data 

```{r}
new<-data.frame(company=1:4,ID=rep(999,4),altid=1:4,choice=rep(T,4))
new$company<-factor(new$company,labels=c("ABB","General Electric","Westinghouse","McGraw-Edison"),ordered=F)
new$company<-relevel(new$company,ref="McGraw-Edison")
#make sure the contrasts are the same as that original data 
contrasts(abb$company)
contrasts(new$company)
```

```{r}
contrasts(abb$company) == contrasts(new$company)
```

```{r}
newld <- mlogit.data(data=new, choice = "choice", 
                  shape = "long",chid.var="ID",
                  alt.var="altid")
cbind(new,predict(fit,newld))
```

### Compare with `prop.table()`

```{r}
#compare predicted share with frequency table
tb<-table(abb$company,abb$choice)
prop.table(tb,margin=1)
```

Gives the same result if only using 1 variable for multinomial logit


## Part (b)

```{r}
fit<-mlogit(formula=choice~0+company+price+energyloss+
              maintenance+warranty+spareparts+installation+
              problem+quality,data=ld,
            alt.var="altid",
            shape="long")
summary(fit)
```

## Part (c)

```{r}
#all attributes are in 9-point scale
range<-abs(fit$coefficients[4:11])*8
range/sum(range)
```

## Part (d)

```{r}
new<-read.table(text=
"4,5,6,1,3,3,3,7,1,999,1
2,2,7,2,5,2,6,4,2,999,2
7,3,4,5,1,1,6,4,3,999,3
2,4,3,1,3,5,4,7,4,999,4",sep=',',header=F)

names(new)<-names(abb[c(5:12,16,1,19)])
new$company<-factor(new$company,labels=c("ABB","General Electric","Westinghouse","McGraw-Edison"))
new$company<-relevel(new$company,ref="McGraw-Edison")
new$choice<-T

newld <- mlogit.data(data=new, choice = "choice", 
                  shape = "long",chid.var="ID",
                  alt.var="altid")

cbind(new,pred=predict(fit,newdata=newld))

```

