---
title: "Tutorial 2"
output: 
  html_notebook:
    toc: true
---

# Tutorial 2

# Libraries

```{r, warning=FALSE}
library(dplyr)
library(tidyverse)
library(psych)
library(magrittr)
library(reshape2)
```


# Question 3

The survey contains the results from 20 customers on broadband internet service. 

* The customers’ income X1
* Age X2 
* Education X3 (in years)
* Expenses on internet service X4 (per month)
* Hours of service X5 (per day)
* the type of internet they used Y 
  * (1-PCCW Ltd.; 2-New World Telecommunications Ltd.; 3-Pacific Internet (HK) Ltd.)

## EDA

```{r}
internet <- read.csv("~/MSTAT/STAT8019_Marketing_Analytics/data/TUT0301.csv")
str(internet)
head(internet)
```

```{r}
describe(internet, skew = F, ranges = F)
```

### Distribution Plot

```{r}
internet%>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density()  
```

## Hierarchicul Clustering

### Standardization

```{r}
# Use standard scaling on variables
# exclude Y , hence only 5 variables to cluster
internet_sc <- scale(internet[, 2:6]) ; internet_sc


```

### Proximity Matrix (using Squared Euclidean Distance)

```{r}
internet_dist <- dist(internet_sc, method = "euclidean")^2
round((internet_dist), 3)
```

### Ward's Method as Distance Between Clusters

```{r}
internet_fit <- hclust(internet_dist, method = "ward.D")
summary(internet_fit)
```

#### Merging History

```{r}
# get merging history
internet_history <- cbind(internet_fit$merge, internet_fit$height)
internet_history
```

### Selecting Appropriate Number of Clusters

#### Distance plot 

```{r}
ggplot(mapping=aes(x=1:length(internet_fit$height),y=internet_fit$height))+
  geom_line()+
  geom_point()+
  labs(x="stage",y="height") + ggtitle("Displot Plot")
```

* The elbow above is at 3 clusters, hence 3 clusters are optimum and suggested

#### Dendogram

```{r}
par(mar=c(1,4,1,3))
plot(internet_fit,labels=internet$case,hang=-1,main="")
axis(side = 2, at = seq(0, 100, 10))
```

### Cluster Analysis

```{r}
internet_clusters <- cutree(internet_fit, k=3)
internet_case <- internet$case
internet_sol <- data.frame(internet_clusters, internet_sc, internet_case)
internet_sol[order(internet_sol$internet_clusters), c("internet_clusters", "internet_case")]
```

#### Interpreting Cluster Characteristics

```{r}
#cluster means
tb<-aggregate(x=internet_sol[,2:6], by=list(cluster=internet_sol$internet_clusters),FUN=mean)
print(tb,digits=2)
```

```{r}

#profile plot
tbm<-melt(tb,id.vars='cluster')
tbm$cluster<-factor(tbm$cluster)
ggplot(tbm, 
       aes(x = variable, y = value, group = cluster, colour = cluster)) + 
  geom_line(aes(linetype=cluster))+
  geom_point(aes(shape=cluster)) +
  geom_hline(yintercept=0) +
  labs(x=NULL,y="mean")
```

* C2: Low income, younger, people, perhaps fresh grads
* C1: Middle class people and lifestyle
* C3: High income lifestyle

#### Numbers of customers in clusters

```{r}
tbl <- as.data.frame(table(internet_sol$internet_clusters))

colnames(tbl) <- c("Cluster", "Frequency") 

ggplot(tbl, aes(factor(Cluster), Frequency, fill = Cluster)) +     
  geom_col(position = 'dodge')
```

# Question 4

## EDA

```{r}
secondary <- read.csv("~/MSTAT/STAT8019_Marketing_Analytics/data/TUT0302.csv")
str(secondary)
head(secondary)
```


```{r}
describe(secondary, skew = F, ranges = F)
```

### Distribution Plot

```{r}
secondary %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
    facet_wrap(~ key, scales = "free") +   # In separate panels
    geom_density()  
```

### Correlation Plot

```{r}
cormat <- secondary %>% 
  keep(is.numeric) %>%
    cor() %>% 
      round(2)

cormat
```

```{r}
melted_cormat <- melt(cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + scale_fill_continuous(high = "#132B43", low = "#56B1F7")
```

## K-Means Clustering

```{r}
# Standard scaling

secondary_sc <- secondary %>%
  keep(is.numeric) %>%
    scale()

secondary_sc
```


### MacQueen (default)
```{r}
set.seed(12345)

fit_mq <- kmeans(x=secondary_sc, centers = 4, algorithm = "MacQueen")

fit_mq
```

### Hartigan Wong
```{r}
fit_hw <- kmeans(x=secondary_sc, centers = 4, algorithm = "Hartigan-Wong")
fit_hw
```

* K-Means using MacQueen is a better fit based on the BSS/TSS ratio

## Hierarchical Clustering

Use its centroids for K means

```{r}
secondary_dist <- secondary_sc %>%
  dist(method="euclidean") 
    
secondary_dist <- secondary_dist^2

```

```{r}
second_hclust_fit <- hclust(secondary_dist, method = "ward.D")
ward.sol <- cutree(second_hclust_fit, k=4)
```

```{r}
# save the centers of each column
tb<-aggregate(secondary_sc,by=list(ward=ward.sol),FUN=mean)
tb

```

### Feed Centroids into K Means

```{r}
fit_k_w <- kmeans(x=secondary_sc, centers = tb[, 2:ncol(tb)], algorithm = "MacQueen")
fit_k_w
```

* Verify that we get back the same results are the `hclust` using Ward's method

```{r}
fit_k_w$cluster == ward.sol
```

### Cluster Analysis

#### Cluster Members
```{r}
as.data.frame(table(fit_k_w$cluster))
```

* Classes are almost balanced

#### Means of Clusters

```{r}
second_comb <- data.frame(cluster=fit_k_w$cluster, secondary_sc) ; second_comb
```


```{r}
second_tb <- second_comb %>% 
  group_by(cluster) %>%
    summarise_all(list(mean=mean)) ; second_tb
```

#### Line Plot

```{r}
#profile plot
tbm<-melt(second_tb,id.vars='cluster')
tbm$cluster<-factor(tbm$cluster)
ggplot(tbm, 
       aes(x = variable, y = value, group = cluster, colour = cluster)) + 
  geom_line(aes(linetype=cluster))+
  geom_point(aes(shape=cluster)) +
  geom_hline(yintercept=0) +
  labs(x=NULL,y="mean")
```

### Class Characteristics

* C1: Students who are good in other subjects except Math
* C2: Students who are not good in any subjects
* C3: Students who are good in every sunject, especially Math
* C$: Students who are moderate in every subject

